Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers

### 2021-09-08 docs/manuscript/manuscript.Rmd
Remove old abstract, edit data/code section

---
title             : "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
shorttitle        : "Registration-publication links in German clinical trials"
author: 
  - name          : "Maia Salholz-Hillel\\orcidlink{0000-0003-1934-9504}"
    affiliation   : "1"
    corresponding : yes
    address       : "Anna-Louisa-Karsch-Str. 2, 10178, Berlin, Germany"
    email         : "maia.salholz-hillel@charite.de"
    role:         # CRediT roles (https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Software
      - Investigation
      - Data Curation
      - Formal Analysis
      - Visualization
      - Validation
      - Project Administration
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Daniel Strech\\orcidlink{0000-0002-9153-079X}"
    affiliation   : "1"
    role:
      - Funding Acquisition
      - Resources
      - Conceptualization
      - Writing - Review & Editing
      - Supervision
  - name          : "Benjamin Gregory Carlisle\\orcidlink{0000-0001-8975-0649}"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
      - Methodology
      - Software
      - Validation
      - Supervision
affiliation:
  - id            : "1"
    institution   : "QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin"
authornote: >
abstract: |
  *ORIGINAL*

  **Objective** To evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). Inadequate links make trial publications and registrations less findable and compromise evidence synthesis and health policy decision making. The World Health Organization (WHO) and others call for better adoption of this straightforward transparency practice. <!--#The World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) call for transparent registration-publication linking to make clinical trial publications and registrations findable and increase the value of clinical research.-->

  **Design** Cross-sectional bibliographic study

  **Setting** German UMC clinical trials

  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and trial publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.

  **Main outcome measures** The proportion of trial registrations that reference a known results publication. The proportion of results publications that report the known TRN in the metadata, abstract, and full-text.

  **Secondary analyses** We constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.

    **Results** Only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_full")` of trials have a linked publication (DOI or PMID) in the registration as well as the TRN in the publication full-text, abstract, and metadata, and only `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft_abs")` of trials met the CONSORT and ICMJE guidelines to include TRNs in both the full-text and the abstract. `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links. The most common link was TRN reported in the full-text (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_ft")`). ClinicalTrials.gov trials were overall better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

  **Conclusion** German UMCs have not comprehensively linked trial registrations and publications by both including a structured reference to the publication in the registration, and reporting TRNs in results publications. In addition to improved linking by trialists, changes in the PubMed TRN capturing process (such as automated strategies like those developed in this study) and automated indexing of publications in DRKS would make trial registrations and results more findable.

  **Open Data and Code** All code and the final analysis data for this study are available at https://github.com/maia-sh/reg-pub-link. Raw data is available at https://github.com/maia-sh/intovalue-data.

  *CLINICAL TRIALS*

  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and interlink components of the same clinical trial. To support the finding and linking of trial evidence, the World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) ask researchers to provide the trial registration number (TRN) in their publication and a reference to the publication in the registration. This straightforward transparency practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient. The aim of this cross-sectional study is to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with concrete guidance to improve trial links.

  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of all registered, published clinical trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and trial registrations numbers (TRNs) in publication metadata, abstract, and full-text.
  We evaluated the proportion of trial registrations that reference a known results publication, and the proportion of results publications that report the known TRN in the metadata, abstract, and full-text. We also constructed exploratory logistic regression models to investigate the relationship between trial completion date, registry, and registration-publication linking.

  **Results, positive or negative? which values to include?** `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_not_both_ft_abs")` of trials fail to reference TRNs in both the full-text and abstract, in breach of long-standing CONSORT and ICMJE guidelines. `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg")` of trial registrations do not contain links to their results publications. `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so registration and publication could be associated only by manual searching and screening. Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and this difference appears to be driven by PubMed and registry infrastructure, rather than by trialists. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

  **Conclusions** German UMCs have not comprehensively linked trial registrations and publications, despite established recommendations. This shortcoming threatens the quality of evidence synthesis and medical practice, and  burdens researchers to manually search and link trial data. Researchers could easily improve this by copy-and-pasting references between their trial registrations and publications. Other stakeholders could make trial evidence even more findable and interlinked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.

@@

Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> All code, the final analysis data, and a list of all packages used for this study are available at <https://github.com/maia-sh/reg-pub-link>. Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>.

>>>>

---
title             : "Results publications are inadequately linked to trial registrations: An automated pipeline and evaluation of German university medical centers"
shorttitle        : "Registration-publication links in German clinical trials"
author: 
  - name          : "Maia Salholz-Hillel\\orcidlink{0000-0003-1934-9504}"
    affiliation   : "1"
    corresponding : yes
    address       : "Anna-Louisa-Karsch-Str. 2, 10178, Berlin, Germany"
    email         : "maia.salholz-hillel@charite.de"
    role:         # CRediT roles (https://casrai.org/credit/)
      - Conceptualization
      - Methodology
      - Software
      - Investigation
      - Data Curation
      - Formal Analysis
      - Visualization
      - Validation
      - Project Administration
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Daniel Strech\\orcidlink{0000-0002-9153-079X}"
    affiliation   : "1"
    role:
      - Funding Acquisition
      - Resources
      - Conceptualization
      - Writing - Review & Editing
      - Supervision
  - name          : "Benjamin Gregory Carlisle\\orcidlink{0000-0001-8975-0649}"
    affiliation   : "1"
    role:
      - Writing - Review & Editing
      - Methodology
      - Software
      - Validation
      - Supervision
affiliation:
  - id            : "1"
    institution   : "QUEST Center for Responsible Research, Berlin Institute of Health (BIH) at Charité Universitätsmedizin Berlin"
authornote: >
abstract: |
  **Background/Aims** Informed clinical guidance and health policy relies on the ability of clinicians, policymakers, and guideline developers to find comprehensive clinical evidence and link registrations and publications of the same clinical trial together. To support the finding and linking of trial evidence, the World Health Organization (WHO), the International Committee of Medical Journal Editors (ICMJE), and the Consolidated Standards of Reporting Trials (CONSORT) ask researchers to provide the trial registration number (TRN) in their publication and a reference to the publication in the registration. This straightforward transparency practice costs researchers minimal effort and makes evidence synthesis more thorough and efficient. This cross-sectional study aims to evaluate links between registration and publication across clinical trials conducted by German university medical centers (UMCs) and registered in either ClinicalTrials.gov or the German Clinical Trials Registry (DRKS). Secondary aims are to develop an automated pipeline that can be applied to other cohorts of trial registrations and publications, and to provide stakeholders, from trialists to registries, with guidance to improve trial links.

  **Methods** We used automated strategies to download and extract data from trial registries, PubMed, and results publications for a cohort of registered, published trials conducted across German UMCs and completed between 2009 and 2017. We implemented regular expressions to detect and classify publication identifiers (DOI and PMID) in registrations, and TRNs in publication metadata, abstract, and full-text.

  **Results** In breach of long-standing CONSORT and ICMJE guidelines, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_not_both_ft_abs")` of trials failed to reference TRNs in both the abstract and full-text of their results publications. Furthermore, `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg")` of trial registrations did not contain links to their results publications. `r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_none")` of trials had no links, so associating registration and publication required manual searching and screening. Overall, ClinicalTrials.gov trials were better linked than DRKS trials, and PubMed and registry infrastructures appear to drive this difference. Of trials reporting a TRN in the abstract, trials in ClinicalTrials.gov were more likely than trials in DRKS to have the TRN captured in the PubMed metadata. Most (`r knitr::load_cache(label = "abstract-vars", object = "abs_n_trials_link_reg_ctgov_auto")`) ClinicalTrials.gov registrations with a publication link were automatically indexed from PubMed metadata, which is not possible in DRKS.

  **Conclusions** German UMCs have not comprehensively linked trial registrations and publications, despite established recommendations. This shortcoming threatens the quality of evidence synthesis and medical practice, and burdens researchers to manually search and link trial data. Researchers could easily improve this by copy-and-pasting references between their trial registrations and publications. Other stakeholders could make trial evidence even more findable and linked, e.g., PubMed could capture additional TRNs using automated strategies (like those developed in this study), and DRKS could automatically index publications from PubMed.

@@

Data collection, preparation, and analysis was performed in `r r_citation`<!--#aside from parsing PDFs to XML in Python-->.<!--#Additional packages and libraries are specified in the methods.--> Code to recreate the analysis dataset, rerun the analysis, and generate this manuscript is available at <https://github.com/maia-sh/reg-pub-link>. Raw data (with the exception of the full-text of publications) is available at <https://github.com/maia-sh/intovalue-data>. A STROBE checklist for reporting cross-sectional studies in provided in the Supplement.
